---
title: "CNV_Ascertainment_UK_Biobank"
author: "Matthew Oetjens"
date: "10/22/2021"
output:
  html_document: default
  pdf_document: default
---

This project is designed to compare the prevalence of dels and dups at pathogenic
CNV regions. Explore the causes of the divergence in CNV prevalence. For example, does a
diverence in prevalence correlate with a larger IQ delta between the CNVs.

```{r include=FALSE}
################################################################################
## IMPORT LIBRARIES
################################################################################

library(tidyverse)
library(gt)

################################################################################
## SETWD AND LOAD DATA
################################################################################

# set working directory in my storage
setwd("/nas/Oetjens_Group_t1/mtoetjens/cnv_ascertainment/")

# load data
ukb_cnv_table.dat <- read.csv("/nas/Oetjens_Group_t1/mtoetjens/cnv_ascertainment/data/CRAWFORD_UKB_CNV_PREVALENCE.csv")
ghs_cnv_table.dat <-read.csv("/nas/Oetjens_Group_t1/mtoetjens/cnv_ascertainment/data/GESINGER_CNV_PREVALENCE.csv")
kendall_cognitive_results.dat <-read.csv("/nas/Oetjens_Group_t1/mtoetjens/cnv_ascertainment/data/KENDALL_COGNITIVE_SCORE_RESULTS.txt")


# convert to tibble
ukb_cnv_table.dat <- as_tibble(ukb_cnv_table.dat) 
ghs_cnv_table.dat <- as_tibble(ghs_cnv_table.dat) 


################################################################################
## Rename CNVs named after syndromes
################################################################################


################################################################################
## UKB CNV Prevalences
################################################################################

# Sample size of CNV table from Crawford et al.
cel_files = 488415
exclusions = 25069
samplesize = cel_files-exclusions

```

```{r include=TRUE}
# Calculate CNV prevalence differences 
ukb_prev_table.dat <- ukb_cnv_table.dat %>%
  dplyr::select(Dosage, CNV, N_carriers) %>%
  mutate_at("CNV", str_replace, "_?del|_?dup", "") %>%
  pivot_wider(names_from = Dosage, values_from =  N_carriers) %>%
  #pivot_wider(names_from = Dosage, values_from =  N_carriers, values_fill=0) %>%
  rename(Del_ct = DEL, Dup_ct = DUP) %>%
  mutate(CNV_ct = Del_ct + Dup_ct, Del_prev_pct = Del_ct/samplesize * 100, Dup_prev_pct = Dup_ct/samplesize * 100, Prev_pct_diff = Dup_prev_pct - Del_prev_pct, Prev_ratio = Dup_ct/Del_ct)  %>%
  filter(CNV_ct > 10) %>%
  arrange(-Prev_ratio)
```

```{r include=TRUE}
################################################################################
## GHS CNV Prevalences
################################################################################
ghs_samplesize = 145000

ghs_prev_table.dat <- ghs_cnv_table.dat %>%
  rename(CNV = Crawford_CNV, N_carriers = count, Dosage=Pathogenic_Dosage) %>%
  mutate_at("CNV", str_replace, "_?del|_?dup", "")  %>%
  pivot_wider(names_from = Dosage, values_from =  N_carriers)  %>%
  rename(Del_ct = DEL, Dup_ct = DUP) %>%
  mutate(CNV_ct = Del_ct + Dup_ct, Del_prev_pct = Del_ct/samplesize * 100, Dup_prev_pct = Dup_ct/samplesize * 100, Prev_pct_diff = Dup_prev_pct - Del_prev_pct, Prev_ratio = Dup_ct/Del_ct)  %>%
  filter(CNV_ct > 10) %>%
  arrange(-Prev_ratio) 
```

tmp.dat <- merge(ghs_prev_table.dat, ukb_prev_table.dat, by ="CNV")
tmp.dat <- tmp.dat[which(tmp.dat$CNV != "15q13.3(CHRNA7)"),]
cor.test(tmp.dat$Prev_ratio.y, tmp.dat$Prev_ratio.x)





```{r}
plot(cars)
```

